### 2017-04-30 jedm-reviews/round-1/reviewer_c_response.md

We have attempted to modify the paper to be more precise about our claims
about what is automatically discovered and what is manually discovered. We
still feel, however, that this method is useful in the sense that these
patterns cannot be discovered from the raw clickstream data manually by
instructors without the application of a model that can provide statistical
summaries of the data such as the ones provided by our model, and no
existing model can be used directly to generate the same behavior patterns
as what our model generates.

> Moreover, the discoveries do not go beyond obvious findings like
> “watching video” states and “reading forum thread” states.

While it is indeed sometimes challenging to discover what the meaning of a
state is, we argue that there are more non-obvious findings later in the
section, where we discover that a student's probability of being in one
particular state that indicates more diverse activity when taking quizzes
correlates negatively with grade outcomes.

A more important point is that the proposed model is a general model that can be applied to any such student interaction data. It is reasonable to expect that the model would be able to generate even more interesting patterns when it is applied to a data set with such interesting patterns. The specific data set used in our experiments is only to show the capacity of the model to reveal interesting and meaningful patterns in the data.

### 2017-04-30 jedm-reviews/round-1/reviewer_c_response.md

We have attempted to modify the paper to be more precise about our claims
about what is automatically discovered and what is manually discovered. We
still feel, however, that this method is useful in the sense that these
patterns cannot be discovered from the raw clickstream data manually by
instructors without the application of a model that can provide statistical
summaries of the data such as the ones provided by our model, and no
existing model can be used directly to generate the same behavior patterns
as what our model generates.

> Moreover, the discoveries do not go beyond obvious findings like
> “watching video” states and “reading forum thread” states.

While it is indeed sometimes challenging to discover what the meaning of a
state is, we argue that there are more non-obvious findings later in the
section, where we discover that a student's probability of being in one
particular state that indicates more diverse activity when taking quizzes
correlates negatively with grade outcomes.

A more important point is that the proposed model is completely general and thus can be applied to analyze any other data set. The actual utility of the patterns discovered from a particular data set is inevitably limited by the true patterns that the data set actually contains; it is reasonable to expect the model to be able to reveal even more interesting patterns when applied to data sets that contain such patterns.

@@

> It’s hard to say this from reviewing the figures alone, but the states
> seem to be characterized almost entirely by the self-loops in the Markov
> chain. If that is the case, then surely simpler averaging methods are
> reasonable.

It is indeed the case that the latent state transitions are _mostly_
described by self-loops, but we make a claim in table 2 that the latent
state transition probabilities are also important in that the "low"
performing students tend to transition out of state 2 more than the
"perfect" performing students (essentially saying that students that get
perfect quiz scores are more likely to continue to participate heavily in
the forum between browsing sessions compared to students that performed
poorly on quizzes). You can also see this visibly if you compare Figure
5(b) and Figure 5(c)'s edge weight from transitioning from state 2 to state
1, which is significantly darker in 5(c) than 5(b). Moreover, although the magnitude of probabilities of the transitions is small, relative comparison of these probabilities across different "contexts" (e.g., across different student groups or different periods of a course) may reveal sufficiently significant differences to understand how student behavior varies according to the context.

As already mentioned earlier, the specific data set used in our experiments is only to show the potential of the model to reveal interesting and meaningful patterns in the data; the intended main contribution of the paper lies in the novel representation of behavior patterns and the novel algorithm to extract such patterns automatically from data, which would empower all instructors to potentially turn any such interaction data into useful insights and knowledge for improving instruction, as well as enable a novel way of performing data-drive research on student behavior (e.g., by doing large scale analysis of the extracted behavior patterns in association with other "context" variables as mentioned above to reveal detailed understanding of student behavior in general).

### 2017-04-30 sec-limitations.tex
Further polished wording about the multiple local maxima where we claimed that the strong patterns tend to always show up.

Our model fits its parameters using maximum likelihood estimation using the
EM algorithm. The EM algorithm is a hill climbing algorithm that is
optimizing in a highly non-convex parameter space. Thus, it can only
guarantee reaching a local maximum in practice~\citep{Dempster:1977:JRSS}.
This may mean that the parameters found for the model may not be the
``best'' parameters in a global sense, which may lead to suboptimal latent
state representations and transition patterns. The problem may not be as serious as it 
appears to be because strong patterns tend to always show up as long as the algorithm reaches a reasonably good local maximum, and the differences of the results tend to be 
related to weak ``unstable'' patterns which may not be reliable anyway. 
 A commonly applied approach
to address the problem of multiple local maxima is to run the model multiple times with different
starting points to allow the model to explore a larger portion of the
potential parameter space. One can then compare the log-likelihood of the
data between the models that were started from different initialization
points and select the one that has the highest value. This still does not
guarantee finding a global maximum, but it does help alleviate the
potential for finding a particularly bad local maximum.  
In practice, we
have found that our model can be fit to the data fairly quickly on
commodity hardware, so running it multiple times to address this concern is
not computationally unfeasible.

>>>>

Our model fits its parameters using maximum likelihood estimation using the
EM algorithm. The EM algorithm is a hill climbing algorithm that is
optimizing in a highly non-convex parameter space. Thus, it can only
guarantee reaching a local maximum in practice~\citep{Dempster:1977:JRSS}.
This may mean that the parameters found for the model may not be the
``best'' parameters in a global sense, which may lead to suboptimal latent
state representations and transition patterns. Empirically, however, we observed
in our experiments that strong patterns tend to always show up as long as the algorithm reaches a reasonably good local maximum, and the differences of the results tend to be 
related to weak ``unstable'' patterns which may not be reliable anyway. Since
it is far more important and useful to reveal the strong salient patterns than weak unreliable ones for instructors, the problem might not necessarily affect the utility of the approach so significantly.  
 A commonly applied approach
to address the problem of multiple local maxima is to run the model multiple times with different
starting points to allow the model to explore a larger portion of the
potential parameter space. One can then compare the log-likelihood of the
data between the models that were started from different initialization
points and select the one that has the highest value. This still does not
guarantee finding a global maximum, but it does help alleviate the
potential for finding a particularly bad local maximum.  
In practice, we
have found that our model can be fit to the data fairly quickly on
commodity hardware, so running it multiple times to address this concern is
not computationally unfeasible.
